A {practical} review on linear and nonlinear global approaches to stability analysis

### 2018-07-17 99_Documentation/ARTICLE_STABFEM/ARTICLE_ASME_V3diogo.tex

%%% use twocolumn and 10pt options with the asme2ej format
\documentclass[twocolumn,10pt]{asme2ej}

\usepackage{epsfig} %% for loading postscript figures
\usepackage{amsmath}

@@

We start from the general problem of a flow field $[{\bf u},p]$ satisfying the incompressible Navier-Stokes equations on a domain $\Omega$,
\begin{eqnarray} \label{NSprimitive}
\partial_t {\bf u} = {\cal NS} ({\bf u};p)
\equiv - {\bf u} \cdot \nabla {\bf u} - \nabla p + \frac{2}{Re}  \nabla \cdot {\mathsf{D}}({\bf u})  \\
\nabla \cdot {\bf u} = 0,
\end{eqnarray}

with suitable boundary conditions on the frontier $\partial \Omega$ of the domain.
Here $ {\mathsf{D}}({\bf u}) $ is the rate-of-strain tensor defined as
$$
 {\mathsf{D}}({\bf u}) = 1/2
\left( \nabla {\bf u} +  \nabla^T  {\bf u}\right)
$$ 
In the framework of finite element methods, we need to write the equation in weak form.
Prior to this we define a scalar product as follows, for either scalar or vectorial quantities 
$\left< \phi_1, \phi_2 \right> $:
$$
\left< \phi_1, \phi_2 \right> = \int_\Omega \overline{\phi}_1 \cdot \phi_2   \mbox{ d} \Omega
$$
The weak form of the Navier-Stokes equations is readily defined by introducing test functions 
$[{\bf v},q]$ associated with the momentum and continuity equations, and integrating over the domain\footnote{In the simple presentation given here we have omitted the issue of boundary conditions. Details on way boundary conditions can be incorporated in the weak formulation through integration by parts can be found in the appendix.}

\be{NSweak}
\forall [{\bf v},q], \quad \partial_t \left< {\bf v}, {\bf u}\right> = \left< {\bf v} , {\cal NS} ({\bf u};p) \right> + \left< q, \nabla \cdot {\bf u}\right>.
\ee

@@

We look for a steady base-flow $({\bf u}_b;p_b)$ satisfying the steady Navier-Stokes equations, i.e. 
${\cal NS} ({\bf u}_b,p_b) = 0$.
Suppose that we have a 'guess' for the base flow $[{\bf u}_b^g,p_b^g]$  which almost satisfies the equations.  We look for a better approximation under the form
\be{Newton1}
[{\bf u}_b,p_b]  = [{\bf u}_b^g,p_b^g] + [\delta {\bf u}_b, \delta p_b].
\ee
Injecting (\ref{Newton1}) into the weak form (\ref{NSweak}) of the Navier-Stokes equations and developing up to linear terms in terms of the perturbation lead to  
${\cal NS}  ({\bf u}_b^g,p_b^g) + {\cal L} _{{\bf u}_b^g}(\delta {\bf u}_b,\delta p_b;{\bf u}_b^g)$, which can also be written in weak form as :
%\be
%{\cal C}( {\bf u}_b^g , \delta {\bf u}_b) -\nabla \delta p_b 
%-2 \nu \nabla  {\bf D}(\delta {\bf u}_b)
%+
%\left[{\bf u}_b^g \cdot \nabla {\bf u}_b^g - \nabla p_b^g -2 \nu {\bf D}(\delta {\bf u}_b^g) \right]
% \qquad \mbox{with} \nabla \delta {\bf u} = 0. 
%\ee
%Or equivalently, in weak form :
\bes{Newton2}
&\left< {\bf v}, {\cal NS} ({\bf u}_b^g)\right> + \left<q, \nabla \cdot {\bf u}_b^g\right>  
\\
+ &\left< { \bf v}, {\cal L} _{{\bf u}_b^g}( \delta {\bf u}_b,\delta p_b; {\bf u}_b^g) \right> + \left<q, \nabla \cdot \delta{\bf u}_b\right> = 0,
\ees
where ${\cal L}$ is the linearised Navier-Stokes operator, defined by its action on a flow field $({\bf u} ; p)$ as follows 
\be{defNSL}
 {\cal L}_{{\bf U}}( {\bf u}, p; {\bf U}) = - {\cal C}( {\bf U} , {\bf u}) -\nabla p
+\frac{2}{Re} \nabla  \cdot {\mathsf{D}}({\bf u}), %\quad ( \mbox{ with } \nabla \cdot {\bf u } = 0 \, ).
 \ee
and ${\cal C}$ is the convection operator defined by 
\be{defC}
{\cal C}( {\bf U} , {\bf u}) = \left( {\bf U} \cdot \nabla \right) {\bf u} + \left( {\bf u} \cdot \nabla \right)  {\bf U}.
\ee

This problem can now be discretized by projecting upon a basis of Taylor-Hood $(u,v,p) \rightarrow (P2,P2,P1)$ finite elements. Noting $\delta X$ the discretization of ($\delta {\bf u}_b;\delta p_b$) this eventually leads to a matricial problem of the form $A \cdot \delta X = Y$. The procedure of Newton iteration is to solve iteratively this set of equations up to convergence.
In our implementation, the algorithm is written in the Freefem++ solver {\em Newton\_2D.edp} 
which is wrapped by the Matlab driver {\em Freefem\_BaseFlow.m}.

@@

\be{startmodepropre}
{\bf u} ({\bf{x}},t) = {\bf u}_b({\bf{x}}) + \epsilon \hat{\bf u}({\bf{x}}) e^{\lambda t}, \,\,\,\, {p}({\bf{x}},t) = {p}_b({\bf{x}}) + \epsilon \hat{p}({\bf{x}}) e^{\lambda t}
\ee
where $\lambda = \sigma + i \omega$ is the eigenvalue, $\sigma$ the amplification rate,
$\omega$ the oscillation rate, $\hat{{\bf u}},\hat{p}$ the eigenmodes, and $\epsilon$ a small parameter.
The eigenmodes and the eigenvalues are the solution of the following eigenproblem:

\be{LEP}
\lambda \hat{{\bf u}} = {\cal L}_{{\bf u}_b}( \hat{\bf u},\hat{p}),
\ee
or, in weak form : 
\be{eigenvalueproblem}
\lambda \left< {\bf v} , \hat{{\bf u}} \right> = \left< {\bf v}, {\cal L}_{{\bf u}_b} ( \hat{\bf u},\hat{p})\right> + \left< q, \nabla \cdot \hat{\bf u} \right>.
\ee
After discretization, we end up with an eigenvalue problem with the matricial form
\be{Eigen_matricial}
\lambda {\mathsf{B}} \hat{X} = {\mathsf{A}} \hat{X}
\ee
Where ${\mathsf{A}}$ is the matrix resulting from the discretization of ${\cal L}_{{\bf u}_b}$, i.e. the same matrix  appearing in the Newton computation of the base flow, and  $B$ is a 'weight' matrix associated to the scalar product $\left<v,u\right> = \int \overline{\bf v} \cdot {\bf u} \mbox{ d} \Omega$.

@@

\be{drag}
D = F_x ({\bf u},p) = 
 \int_{\Gamma_{cil}} \left[-p {\bf n} + \frac{2}{Re} {\bf D}({\bf u}) \cdot {\bf n} \right]   \cdot {\bf e_x} d\ell
\ee

@@

%For $Re\right>Re_c$, the linear stability analysis based on the {\em base flow} does not allow to predict the properties of the limit cycle; in particular the Strouhal number is substantially underestimated. %It has been shown that investigating the stability properties of the {\em mean flow} obtained by averaging time-marching solution gives a better accordance. Attempts to on a rigourous basis  
In the past decade, efforts have been devoted to extend the range of validity of global approaches into the nonlinear regime for $Re>Re_c$, with the double objective to describe the properties of the limit cycle reached after saturation and to derive amplitude equations describing the transitent dynamics towards this cycle. The two main milestones in this direction is the {\em weakly nonlinear model}  (WNL) of Sipp \& Lebedev \cite{SippLebedev} and the {\em self-consistent model} (SC) of Manti\v{c}-Lugo et al \cite{MLugo2014}.
 {\color{blue} In this review, we will address only the first of the two questions, namely description of the saturated cycle, and leave aside the question of transient dynamics. We will successively review the two aforementioned models, in a simplified formulation devoted to describe only the saturated cycle. We also provide a simple implementation of these two models. Finally, in line with the previous section, we will again illustrate with results obtained for the wake of a cylinder.
 }

\paragraph{General definitions in the nonlinear régime}

In this part, to simplify the notations, we symbolically write the Navier-Stokes equations, as $\partial_t {\bf u} = {\cal 
NS} ({\bf u};p)$, therefore dropping the systematic reference to the 
incompressibility constraint and associated pressure field.

In the nonlinear régime, the {\em base flow} introduced in the linear theory is not directly accessible and no longer relevant, especially when the oscillation amplitudes become large. 
Instead, we will define a {\em mean flow}  as the time-average  of the flow :
\be{defumean}
{\bf u}_m({\bf x})  = \frac{1}{T} \int_0^{T}  {\bf u}({\bf x},t)  dt
\ee
where $T = 2\pi/\omega$ is the period of the oscillation cycle.

A convenient measure of the unsteady part of the flow, which has been adopted in both the WNL and SC models, is the energy-amplitude, defined as the square-root of the total energy associated to the fluctuating part of the velocity field :

\be{defAnergy}
A_E = \sqrt{\frac{1}{T} \int_0^{T} \left( \int_\Omega | {\bf u} - {\bf u}_m |^2 \, dS \right) dt}
\ee

Another quantity which has been particularly addressed in previous studies is the mean recirculation length $L_x$
already discussed introduced and documented in figure ??. In the nonlinear regime, the recirculation bubble becomes unsteady so it is difficult to define properly an instantaneous recirculation length. It is still possible to define a mean recirculation length associated to the mean flow.

In the next sections we will document the predictions of the WNL and SC models regarding the quantities $A_E$ and $L_x$ adopted in past studies. In addition, we will document two other quantities of practical interest : the drag and Lift forces $F_x$ and $F_y$ exerted on the cylinder. 
Both are periodic functions of time, and owing to symmetry consideration, the drag contains only even harmonics and the lift only odd harmonics :
$$
F_x  = F_{x,0}  + \Sigma \left(  F_{x,2n,c} \cos 2 n \omega t + F_{x,2n,s} \sin 2 n  \omega t \right)
$$
$$
F_y  =  \Sigma \left(  F_{y,{2n+1},c} \cos (2n+1) \omega t + F_{y,(2n+1),s} \sin (2n+1) \omega t \right)
$$

In the sequel we will focus on the {\em mean drag}
 $F_{x,0}$ and the part of the fundamental frequency part of the lift  given by $F_{y,1,c}$ and $F_{y,1,s}$.

@@

\paragraph{Weakly nonlinear model}

\begin{figure*}
\small
\begin{lstlisting}
[bf,em]=SF_FindThreshold(bf,em);
wnl = SF_WNL(bf,em);

@@

for Re = Re_HB(2:end)
    [meanflow,mode] = SF_HarmonicBalance(meanflow,mode,'Re',Re);
    omega_HB = [omega_HB imag(mode.lambda)];
    Aenergy_HB  = [Aenergy_HB mode.Energy];
end

\end{lstlisting}
\normalsize
\caption{Illustration of the procedure for nonlinear calculations using StabFem (from script {\em SCRIPT\_CYLINDER.m}. }
\label{fig:listingNL}
\end{figure*}

We give here a brief account of the weakly nonlinear model of Sipp \& 
Lebedev\cite{SippLebedev}, also discussed by Gallaire et al. \cite{FDR2016}. 
The derivation is based on a multiple scale  decomposition based on the parameter $\epsilon^2 = 1/Re - 1/Re_c$. 
In the case where we are interested only in the desciption of the saturation cycle, the starting point can be taken as the following expansion of the state-vector ${\bf{q}}=({\bf{u}},p)$ defining the flow :

\begin{eqnarray}
{\bf q} &=& {\bf q}_{bc} + \epsilon \left[ A_{wnl}  \hat{\bf q} e^{i (\omega_c+\epsilon \omega_\epsilon)  t} + c.c. \right] \label{WNL1}\\
&+& \epsilon^2 \left[ {\bf q}_\epsilon + |A_{wnl}|^2  {\bf q}_{2,0} + \left(  A_{wnl}^2 {\bf q}_{2,2} e^{2 i(\omega_c+\epsilon \omega_\epsilon)  t} + c.c. \right) \right] + {\cal O}( \epsilon ^3)
\nonumber
\label{eq:WNLdevelopment}
\end{eqnarray}
where ${\bf q}_{bc}$ is the base flow at the threshold $Re_c$, $\hat{\bf q}$ the neutral 
eigenmode at $Re_c$, $A_{wnl}$, an amplitude, $\omega_c$ the frequency predicted by the linear approach at $Re=Re_c$ and $\omega_\epsilon$ a small deviation of the frequency.

Note that this starting point is slightly different from the original approach of \cite{SippLebedev} who used a multiple scale method and assumed a dependence of the amplitude $A_{wnl}$ as function of a slow time-scale $\tau = \epsilon t$. This latter starting point is the righth way to parametrize the problem if one wants to obtain an amplitude equation, but as specified above our goal here is to provide a simpler formulation descibing only the saturated cycle.

Details of the approach are given in the appendix. The outcome is a relationship with the form

\be{WNL3}
i \omega_\epsilon A_{wnl} = \Lambda A_{wnl} - (\nu_0+\nu_2)  |A_{wnl}|^2 A_{wnl},
\ee

Where coefficients $\Lambda$, $\nu_0$ and $\nu_2$ are given by eqs. (???)

If one is interested in the amplitude of the limit cycle, we have 
$|A_{wnl}| = \sqrt{\Lambda_r/(\nu_{0,r}+\nu_{2,r})}$.

% and eventually, reintroducing the scaling :
%\be{ANL} 
%|A| =  \epsilon |A_{wnl}| = \frac{\Lambda_r}{\nu_{0,r}+\nu_{2,r} } \sqrt{\frac{1}{Re_c}-\frac{1}{Re}}.
%\ee

The frequency of the limit cycle is given by
\be{omegaWNL} 
\omega_\epsilon =  - \Lambda_i \frac{\nu_{0,i}+\nu_{2,i} }{\nu_{0,r}+\nu_{2,r} } 
\ee

THE MEAN FLOW IS GIVEN BY :

AND THE ASSOCIATED DRAG FORCE IS GIVEN BY :

{\color{blue}
A key issue in the weakly nonlinear expansion is that the definition of the amplitude depends upon a normalization choice 
of the eigenmodes. Several choices are possible. In the litterature three possibilities have been used :
\begin{itemize}
\item First, \cite{SippLebedev} normalized the eigenmode by assuming a specified value to the $y$-component of the velocity at one point, namely :
\be{normsipp}
\hat{u_y}(1,0) = 0.4612
\ee
The advantage of this choice is that the amplitude is then a measure of the energy of the perturbation: 
$\epsilon A_{wnl} \approx A_E$.

\item Secondly, \cite{FDR2016} proposed the following normalisation choice :
\be{normFDR}
\int_\Omega |\hat{\bf u}|^2 d {\bf x}  = \frac{1}{2}
\ee
This directly leads to $\epsilon A_{wnl} = A_E$.

\item Thirdly, following \cite{Fabre2012}, another convenient choice is to normalize the eigenmode with its lift 
force : 
\be{normfabre}
L(\hat{\bf u};\hat{p}) = 1/2
\ee
The advantage of this choice is that the amplitude $\epsilon A_{wnl}$ is directly related to the fundamental lift :
$
F_{y,1,c} 
%= \epsilon A \left( 1/2 e^{i \omega t} + c.c. \right)
=   \epsilon A \cos (\omega t) ; \quad F_{y,1,c} = 0. 
$

>>>>

%%% use twocolumn and 10pt options with the asme2ej format

\documentclass[twocolumn,10pt]{asme2ej}

\usepackage[none]{hyphenat}
\usepackage{epsfig} %% for loading postscript figures
\usepackage{amsmath}

\usepackage{breqn}

@@

We start from the general problem of a flow field $[{\bf u},p]$ satisfying the incompressible Navier-Stokes equations on a domain $\Omega$,
\begin{eqnarray} \label{NSprimitive}
\partial_t {\bf u} = {\color{red}{\cal NS} ({\bf u},p)}
\equiv - {\bf u} \cdot \nabla {\bf u} - \nabla p + \frac{2}{Re}  \nabla \cdot {\mathsf{D}}({\bf u}),  \\
\nabla \cdot {\bf u} = 0,
\end{eqnarray}

with suitable boundary conditions on the frontier $\partial \Omega$ of the domain.
Here $ {\mathsf{D}}({\bf u}) $ is the rate-of-strain tensor defined as
$$
 {\mathsf{D}}({\bf u}) = 1/2
\left( \nabla {\bf u} +  \nabla^T  {\bf u}\right),
$$ 
In the framework of finite element methods, we need to write the equation in weak form.
Prior to this we define a scalar product as follows, for either scalar or vectorial quantities 
$\left< \phi_1, \phi_2 \right> $:
$$
\left< \phi_1, \phi_2 \right> = \int_\Omega \overline{\phi}_1 \cdot \phi_2   \mbox{ d} \Omega,
$$
The weak form of the Navier-Stokes equations is readily defined by introducing test functions 
$[{\bf v},q]$ associated with the momentum and continuity equations, and integrating over the domain\footnote{In the simple presentation given here we have omitted the issue of boundary conditions. Details on way boundary conditions can be incorporated in the weak formulation through integration by parts can be found in the appendix.}

\be{NSweak}
\forall [{\bf v},q], \quad \partial_t \left< {\bf v}, {\bf u}\right> = \left< {\bf v} , {\color{red}{\cal NS} ({\bf u},p)} \right> + \left< q, \nabla \cdot {\bf u}\right>.
\ee

@@

We look for a steady base-flow {\color{red}$[{\bf u}_b,p_b]$} satisfying the steady Navier-Stokes equations, i.e. 
%${\cal NS} ({\bf u}_b,p_b) = 0$.
{\color{red}${\cal NS} ({\bf u}_b,p_b) = 0$.}
Suppose that we have a {\color{red}``guess''}  for the base flow $[{\bf u}_b^g,p_b^g]$  which almost satisfies the equations.  We look for a better approximation under the form
\be{Newton1}
[{\bf u}_b,p_b]  = [{\bf u}_b^g,p_b^g] + [\delta {\bf u}_b, \delta p_b].
\ee
Injecting (\ref{Newton1}) into the weak form (\ref{NSweak}) of the Navier-Stokes equations and developing up to linear terms in terms of the perturbation lead to  
${\cal NS}  ({\bf u}_b^g,p_b^g) + {\color{red} {\cal L} _{{\bf u}_b^g}(\delta {\bf u}_b,\delta p_b)}$, which can also be written in weak form as :
%\be
%{\cal C}( {\bf u}_b^g , \delta {\bf u}_b) -\nabla \delta p_b 
%-2 \nu \nabla  {\bf D}(\delta {\bf u}_b)
%+
%\left[{\bf u}_b^g \cdot \nabla {\bf u}_b^g - \nabla p_b^g -2 \nu {\bf D}(\delta {\bf u}_b^g) \right]
% \qquad \mbox{with} \nabla \delta {\bf u} = 0. 
%\ee
%Or equivalently, in weak form :
\bes{Newton2}
&\left< {\bf v}, {\color{red}{\cal NS} ({\bf u}_b^g,p_b^g)}\right> + \left<q, \nabla \cdot {\bf u}_b^g\right>  
\\
+ &\left< { \bf v}, {\color{red}{\cal L} _{{\bf u}_b^g}( \delta {\bf u}_b,\delta p_b) }\right> + \left<q, \nabla \cdot \delta{\bf u}_b\right> = 0,
\ees
where ${\cal L}$ is the linearised Navier-Stokes operator, defined by its action on a flow field {\color{red}$[{\bf u} , p]$} as follows 
\be{defNSL}
{\color{red} {\cal L}_{{\bf U}}( {\bf u}, p)} = - {\cal C}( {\bf U} , {\bf u}) -\nabla p
+\frac{2}{Re} \nabla  \cdot {\mathsf{D}}({\bf u}), %\quad ( \mbox{ with } \nabla \cdot {\bf u } = 0 \, ).
 \ee
and ${\cal C}$ is the convection operator defined by 
\be{defC}
{\cal C}( {\bf U} , {\bf u}) = \left( {\bf U} \cdot \nabla \right) {\bf u} + \left( {\bf u} \cdot \nabla \right)  {\bf U}.
\ee

This problem can now be discretized by projecting upon a basis of Taylor-Hood $(u,v,p) \rightarrow (P2,P2,P1)$ finite elements. Noting $\delta X$ the discretization of {\color{red}$[\delta {\bf u}_b,\delta p_b]$} this eventually leads to a matricial problem of the form $A \cdot \delta X = Y$. The procedure of Newton iteration is to solve iteratively this set of equations up to convergence.
In our implementation, the algorithm is written in the Freefem++ solver {\em Newton\_2D.edp} 
which is wrapped by the Matlab driver {\em Freefem\_BaseFlow.m}.

@@

\be{startmodepropre}
{\bf u} ({\bf{x}},t) = {\bf u}_b({\bf{x}}) + \epsilon \hat{\bf u}({\bf{x}}) e^{\lambda t}, \,\,\,\, {p}({\bf{x}},t) = {p}_b({\bf{x}}) + \epsilon \hat{p}({\bf{x}}) e^{\lambda t},
\ee
where $\lambda = \sigma + i \omega$ is the eigenvalue, $\sigma$ the amplification rate,
$\omega$ the oscillation rate, $\hat{{\bf u}},\hat{p}$ the eigenmodes, and $\epsilon$ a small parameter.
The eigenmodes and the eigenvalues are the solution of the following eigenproblem:

\be{LEP}
\lambda \hat{{\bf u}} = {\cal L}_{{\bf u}_b}( \hat{\bf u},\hat{p}),
\ee
or, in weak form : 
\be{eigenvalueproblem}
\lambda \left< {\bf v} , \hat{{\bf u}} \right> = \left< {\bf v}, {\cal L}_{{\bf u}_b} ( \hat{\bf u},\hat{p})\right> + \left< q, \nabla \cdot \hat{\bf u} \right>.
\ee
After discretization, we end up with an eigenvalue problem with the matricial form
\be{Eigen_matricial}
\lambda {\mathsf{B}} \hat{X} = {\mathsf{A}} \hat{X},
\ee
{\color{red} where} ${\mathsf{A}}$ is the matrix resulting from the discretization of ${\cal L}_{{\bf u}_b}$, i.e. the same matrix  appearing in the Newton computation of the base flow, and  $B$ is a 'weight' matrix associated to the scalar product $\left<v,u\right> = \int \overline{\bf v} \cdot {\bf u} \mbox{ d} \Omega$.

@@

\be{drag}
D = F_x ({\bf u},p) = 
 \int_{\Gamma_{cil}} \left[-p {\bf n} + \frac{2}{Re} {\bf D}({\bf u}) \cdot {\bf n} \right]   \cdot {\bf e_x} d\ell,
\ee

@@

%For $Re\right>Re_c$, the linear stability analysis based on the {\em base flow} does not allow to predict the properties of the limit cycle; in particular the Strouhal number is substantially underestimated. %It has been shown that investigating the stability properties of the {\em mean flow} obtained by averaging time-marching solution gives a better accordance. Attempts to on a rigourous basis  
In the past decade, efforts have been devoted to extend the range of validity of global approaches into the nonlinear regime for $Re>Re_c$, with the double objective to describe the properties of the limit cycle reached after saturation and to derive amplitude equations describing the transitent dynamics towards this cycle. The two main milestones in this direction {\color{red} are} the {\em weakly nonlinear model}  (WNL) of Sipp \& Lebedev \cite{SippLebedev} and the {\em self-consistent model} (SC) of Manti\v{c}-Lugo et al \cite{MLugo2014}.
 {\color{blue} In this review, we will address only the first of the two questions, namely {\color{red} the} description of the saturated cycle, and leave aside the question of transient dynamics. We will successively review the two aforementioned models, in a simplified formulation devoted to describe only the saturated cycle. We also provide a simple implementation of these two models. Finally, in line with the previous section, we will again illustrate with results obtained for the wake of a cylinder.
 }

\paragraph{General definitions in the nonlinear r{\color{red}e}gime}

In this part, to simplify the notations, we symbolically write the Navier-Stokes equations, as {\color{red}$\partial_t {\bf u} = {\cal NS} ({\bf u})$}, therefore dropping the systematic reference to the incompressibility constraint and associated pressure field. {\color{red} The same is done with the linearized operator ${\cal L}_{{\bf U}}( {\bf u})$.}

In the nonlinear r{\color{red}e}gime, the {\em base flow} introduced in the linear theory is not directly accessible and no longer relevant, especially when the oscillation amplitudes become large. 
Instead, {\color{red} we taken the Reynolds decomposition ${\bf u}={\bf u}_m({\bf x})+{\bf u}'({\bf x},t)$, with ${\bf u}'$ the nonlinear perturbation and ${\bf u}_m$ the {\em mean flow} defined as the time-average  of the flow:}
\be{defumean}
{\bf u}_m({\bf x})  = \frac{1}{T} \int_0^{T}  {\bf u}({\bf x},t)  dt,
\ee
where $T = 2\pi/\omega$  is the period of the oscillation cycle.

A convenient measure of the unsteady part of the flow, which has been adopted in both the WNL and SC models, is the energy-amplitude, defined as the square-root of the total energy associated to the fluctuating part of the velocity field:
\be{defAnergy}
A_E = \sqrt{\frac{1}{T} \int_0^{T} \left( \int_\Omega | {\bf u} - {\bf u}_m |^2 \, dS \right) dt}.
\ee

Another quantity which has been particularly addressed in previous studies is the mean recirculation length $L_x$
already discussed{\color{red},} introduced and {\color{red}documented in figure \ref{fig:LxandDrag}}. {\color{red} However, in a} nonlinear regime, the recirculation bubble becomes unsteady, {\color{red} becoming} difficult to define properly an instantaneous recirculation length. It is still possible to define a mean recirculation length associated to the mean flow. %On va le faire ??

In the next sections we will document the predictions of the WNL and SC models regarding the quantities $A_E$ and $L_x$ adopted in past studies. In addition, we will document two other quantities of practical interest : the Drag and Lift forces $F_x$ and $F_y$ exerted on the cylinder. 
Both are periodic functions of time and{\color{red},} owing to symmetry consideration, the drag contains only even harmonics and the lift only odd harmonics:{\color{red}
%\be{drag_lift_def}
%\begin{cases}
%F_x=F_{x,0} + \sum_{n=1}^\infty \big( F_{x,2n,c} \cos ( 2 n \omega t) + F_{x,2n,s} \sin( 2 n  \omega t ) \big) \\
%\begin{split}
%F_y  =  \sum_{n=1}^\infty \big(  F_{y,{2n-1},c} \cos ((2n-1) \omega t ) +\\
% F_{y,(2n-1),s} \sin ((2n-1) \omega t) \big)
%\end{split}
%\end{cases}.
%\ee
\begin{align}
F_x &=F_{x,0} + \sum_{n=1}^\infty \big( F_{x,2n,c} \cos ( 2 n \omega t) + F_{x,2n,s} \sin( 2 n  \omega t ) \big),
\\
\begin{split}
F_y  & =\sum_{n=1}^\infty \big( F_{y,{2n-1},c} \cos ((2n-1) \omega t )\\
             &\qquad + F_{y,(2n-1),s}  \sin ((2n-1) \omega t) \big).
\end{split}
\label{drag_lift_def}
\end{align}

In the sequel we will focus on the {\em mean drag}  $F_{x,0}$ and on the fundamental frequency part of the lift given by $F_{y,1,c}$ and $F_{y,1,s}$.
 }

@@

\paragraph{Weakly nonlinear model}
%REFAIRE
\iffalse
\begin{figure*}
\small
\begin{lstlisting}
[bf,em]=SF_FindThreshold(bf,em);
wnl = SF_WNL(bf,em);

@@

for Re = Re_HB(2:end)
    [meanflow,mode] = SF_HarmonicBalance(meanflow,mode,'Re',Re);
    omega_HB = [omega_HB imag(mode.lambda)];
    Aenergy_HB  = [Aenergy_HB mode.Energy];
end
\end{lstlisting}
\normalsize
\caption{Illustration of the procedure for nonlinear calculations using StabFem (from script {\em SCRIPT\_CYLINDER.m}. }
\label{fig:listingNL}
\end{figure*}
\fi
\begin{figure*}
\small
\begin{lstlisting}
[bf,em]=SF_FindThreshold(bf,em); Omegac=imag(em.lambda);
[ev,em] = SF_Stability(bf,'shift',1i*Omegac,'nev',1,'type','S'); 
[wnl,meanflow,mode] = SF_WNL(bf,em,'Retest',47.,'Normalization','L');

for Re = Re_HB(2:end)
    [meanflow,mode] = SF_HarmonicBalance(meanflow,mode,'Re',Re);
    omega_HB = [omega_HB imag(mode.lambda)];
    Aenergy_HB  = [Aenergy_HB mode.Energy];
end
\end{lstlisting}
\normalsize
\caption{Illustration of the procedure for nonlinear calculations using StabFem (from script {\color{red}{\em SCRIPT\_CYLINDER\_NONLINEAR.m})}. }
\label{fig:listingNL}
\end{figure*}

We give here a brief account of the weakly nonlinear model of Sipp \& 
Lebedev\cite{SippLebedev}, also discussed by Gallaire et al. \cite{FDR2016}. 
The derivation is based on a multiple scale  decomposition based on the parameter {\color{red}$\epsilon^2 = 1/Re_c - 1/Re$}. 
In the case where we are interested only in the {\color{red}description} of the saturation cycle, the starting point can be taken as the following expansion of %the state-vector ${\bf{q}}=({\bf{u}},p)$ defining the flow:
 {\color{red}the velocity flow field:} 
 {\color{red}
\be{WNL1}
\begin{aligned}
&{\bf u} = {}  {\bf u}_{bc} + \epsilon \left[ A_{wnl}  \hat{\bf u} e^{i (\omega_c+\epsilon^{\color{red}2} \omega_\epsilon)  t} + c.c. \right]\\
&+ \epsilon^2 \left[ {\bf u}_\epsilon + |A_{wnl}|^2  {\bf u}_{2,0} + \left(  A_{wnl}^2 {\bf u}_{2,2} e^{2 i(\omega_c+\epsilon^{\color{red}2} \omega_\epsilon)  t} + c.c. \right) \right]\\
& + {\cal O}( \epsilon ^3),
\end{aligned}
\ee
}
where {\color{red}${\bf u}_{bc}$} is the base flow at the threshold $Re_c$, {\color{red} $\hat{\bf u}$} the neutral 
eigenmode at $Re_c$, {\color{red} $A_{wnl}$ an amplitude,} $\omega_c$ the frequency predicted by the linear approach at $Re=Re_c$ and $\omega_\epsilon$ a small deviation of the frequency, {\color{red} ${\bf u}_{\epsilon}$ the base flow modifications related to the increase of $\epsilon$, ${\bf u}_{2,0}$ the flow response to the steady nonlinear interactions of $\hat{\bf u}$ with its conjugate, ${\bf u}_{2,2}$ the flow response to the unsteady nonlinear interactions of $\hat{\bf u}$ with itself and  ${\cal O}( \epsilon ^3)$ the third order terms.   }

Note that this starting point is slightly different from the original approach of \cite{SippLebedev} who used a multiple scale method and assumed a dependence of the amplitude $A_{wnl}$ as function of a slow time-scale $\tau = {\color{red}\epsilon^2} t$. This latter starting point is the {\color{red}right} way to parametrize the problem if one wants to obtain an amplitude equation, but{\color{red},} as specified above{\color{red},} our goal here is to provide a simpler formulation {\color{red}describing} only the saturated cycle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%DETAILS IN APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
{\color{red}
The expansion \ref{WNL1} can be introduced into the Navier-Stokes equations \ref{NSprimitive}. Details at the different $\epsilon^i$ orders are given in the appendix}. The outcome is a relationship with the form
\be{WNL3}
i \omega_\epsilon A_{wnl} = \Lambda A_{wnl} - (\nu_0+\nu_2)  |A_{wnl}|^2 A_{wnl},
\ee
where the coefficients $\Lambda$, $\nu_0$ and $\nu_2$ are given by {\color{red}equations \ref{LAMBDA}, \ref{NU0} and \ref{NU2}, respectively.

We note that $\Lambda$ does not depend on the normalization choice of the eigenmode, discussed later, while $\nu_0$ and $\nu_2$ depend. Hence, if one is interested in the amplitude of the limit cycle, the latter will depend on it and will be given by $|A_{wnl}| = \sqrt{\Lambda_r/(\nu_{0,r}+\nu_{2,r})}$.

Reintroducing the scaling
\be{ANL} 
|A| =  \epsilon |A_{wnl}| =\sqrt{ \frac{\Lambda_r}{\nu_{0,r}+\nu_{2,r} }} \sqrt{\frac{1}{Re_c}-\frac{1}{Re}},
\ee
the frequency of the limit cycle is given by
\be{omegaWNL} 
\omega_{LC}=\omega_c+ \epsilon^2\Lambda_i- (\nu_{0,i}+\nu_{2,i})|A|^2,
\ee

where the subscripts $r$ and $i$ represent the real and imaginary parts.
}
{\color{red}
Recasting equation \ref{WNL1} in the {\em mean flow} definition \ref{defumean}, we obtain
\be{mean_flow_WNL}
{\bf u}_m({\bf x })= {\bf u}_{bc}+\epsilon^2 {\bf u}_\epsilon + |A|^2{\bf u}_{2,0},
\ee
where the first two terms give the base flow approximation at the $Re$ prescribed by $\epsilon$. }

{\color{red} Finally, the associated Drag and Lift are given respectively by:
\be{drag_cos}
\begin{aligned}
F_{x,WNL} = F_{x}({\bf u}_{bc})&+\epsilon^2 \big( 
F_{x}({\bf u}_{\epsilon})- 2\mathsf{D}({\bf u}_{bc}).{\bf n}_x \big)\\
&+|A|^2 F_{x}({\bf u}_{2,0})\\
&+2|A|^2\big(F_{x,r}({\bf u}_{2,2})\cos(2\omega t)
-F_{x,i}({\bf u}_{2,2})\sin(2\omega t) \big),
\end{aligned}
\ee
\be{lift_cos}
\begin{aligned}
F_{y,WNL} =2\big(|A|_r F_{x,r}({\bf \hat{u}})\cos(\omega t)-
|A|_i F_{x,i}({\bf \hat{u}})\sin(\omega t) \big),\\
\end{aligned}
\ee
where $F_{x}$ and $F_{y}$ are given by \ref{drag} and \ref{lift} (with the associated pressure field implicit), respectively . All terms in these equations can be matched with the interesting ones in \ref{drag_lift_def}: $F_{x,0}$ is the first 4 terms of \ref{drag_cos}; $F_{x,2,c}$, $F_{x,2,s}$ are the fifth and sixth terms of \ref{drag_cos}; $F_{y,1,c}$, $F_{y,1,s}$ are the first and second terms of \ref{lift_cos}. We note that the \textit{sin} term is associated to the phase of the forces in a physical way. }
